---
title: "00.0_Setup.Rmd"
author: "Jason Torres"
date: "May 13, 2016"
output: pdf_document
---

The purpose of this script is to determine - for each tissue model - which genes have prediction FDR $\leq 0.05$. All subsequent results will be based on genes \textbf{meeting this threshold}. This is a **model-level** correction that ensures that all genes analyzed meet a minimum standard of prediction quality. The requisite prediction information files are on tarbell:  `/group/im-lab/nas40t2/jason/projects/MetaXcan/db_files/gtex/alpha_1/allResults` 

Load libraries and set paths 
```{r setup, echo=FALSE}

library("data.table")
library("dplyr")
library("reshape")
library("qvalue")
"%&%" <- function(a,b) paste0(a,b)

home.dir <- "/Users/jtorres/Google Drive/Projects/" %&% 
  "MetaXcan/DIAGRAM-T2D-version3/"
model.dir <- "/Volumes/im-lab/nas40t2/jason/projects/" %&% 
  "MetaXcan/db_files/gtex/alpha_1/allResults/"
results.dir <- "/Volumes/im-lab/nas40t2/jason/projects/MetaXcan/results/"
temp.dir <- home.dir %&% "temp/" 
dir.create(temp.dir)


```


Create FDR $\leq 0.05$ temporary files 
```{r fdr_filter, echo=FALSE}

file.vec <- list.files(model.dir)
index1 <- grep("CrossTissue",file.vec)
index2 <- grep("TW",file.vec)
index <- c(index1,index2)
file.vec <- file.vec[index]

fdr_filter <- function(fdr=0.05){
  record.df <- c()
  for (f in file.vec){
    mod <- gsub(".allResults.txt.gz","",f)
    print(mod)
    df <- fread("cat " %&% model.dir %&% f %&% " | zmore")
    pvec <- df$pval
    qval <- qvalue(pvec)$qvalues
    df <- cbind(df,qval)
    sub.df <- filter(df,qval<=fdr) %>% select(one_of("gene","genename"))
    out.name <- temp.dir %&% mod %&% ".fdr" %&% fdr %&% ".temp.txt"
    write.table(sub.df,file=out.name,quote=FALSE,sep="\t",row.names=FALSE)
    total <- dim(df)[1]; retain <- dim(sub.df)[1]; prop <- retain/total
    vec <- c(mod,total,retain,prop)
    record.df <- rbind(record.df,vec)
  }
  record.df <- as.data.frame(record.df,stringsAsFactors=FALSE)
  names(record.df) <- c("model","total","retained","prop")
  record.df$model <- as.character(record.df$model)
  record.df$total <- as.numeric(record.df$total)
  record.df$retained <- as.numeric(record.df$retained)
  record.df$prop <- as.numeric(record.df$prop)
  
  rec.name <- temp.dir %&% "model-record-fdr" %&% fdr %&% ".txt"
  write.table(record.df,file=rec.name,quote=FALSE,sep="\t",row.names=FALSE)
  print(hist(record.df$prop,breaks=10))
}

```


Subset \textbf{MetaXcan} results (on tarbell) to include only genes that meet the FDR $\leq 0.05$ threshold above in each respective tissue. 
```{r subset_write, echo=FALSE}

model.vec <- gsub(".allResults.txt.gz","",file.vec)

subset_write <- function(fdr=0.05,analysis="DIAGRAM3"){
  #analysis can be DIAGRAM3 or GERA-T2D
  res1.dir <- results.dir %&% analysis %&% "/alpha_1/"
  res2.dir <- results.dir %&% analysis %&% "/alpha_1_predictFDR" %&% fdr %&% "/"
  dir.create(res2.dir)
  for (mod in model.vec){
    print(mod)
    df1 <- fread(res1.dir %&% mod %&% ".zscores.csv")
    temp.file <- temp.dir %&% mod %&% ".fdr" %&% fdr %&% ".temp.txt"
    df2 <- fread(temp.file)
    df3 <- filter(df1,gene_name %in% df2$genename)
    write.table(df3,file=res2.dir %&% mod %&% ".zscores.csv",sep=",",quote=FALSE,row.names=FALSE)
    unlink(temp.file)
  }
}

fdr_filter(0.05)
fdr_filter(0.10)
subset_write(fdr=0.05,analysis="DIAGRAM3")
subset_write(fdr=0.10,analysis="DIAGRAM3")
fdr_filter(0.05)
fdr_filter(0.10)
subset_write(fdr=0.05,analysis="GERA-T2D")
subset_write(fdr=0.10,analysis="GERA-T2D")
```



Subset \textbf{MetaXcan} results (on tarbell) to include only genes that meet the FDR $\leq 0.05$ threshold above in each respective tissue - this is for the **META-ANALYSIS OF DIAGRAM AND GERA**  
```{r subset_write_meta, echo=FALSE}

subset_write_meta <- function(fdr=0.05){
  res1.dir <- results.dir %&%  "meta-analyses/DIAGRAM-GERA/alpha_1/"
  res2.dir <- results.dir %&%  "meta-analyses/DIAGRAM-GERA/alpha_1_predictFDR" %&% fdr %&% "/"
  dir.create(res2.dir)
  for (mod in model.vec){
    print(mod)
    df1 <- fread(res1.dir %&% mod %&% ".meta.txt")
    temp.file <- temp.dir %&% mod %&% ".fdr" %&% fdr %&% ".temp.txt"
    df2 <- fread(temp.file)
    df3 <- filter(df1,Gene %in% df2$genename)
    write.table(df3,file=res2.dir %&% mod %&% ".meta.txt",sep="\t",quote=FALSE,row.names=FALSE)
    unlink(temp.file)
  }
}

fdr_filter(0.05)
fdr_filter(0.10)
subset_write_meta(fdr=0.05)
subset_write_meta(fdr=0.10)
```


