---
title: "Eval-modSNPs-ImpG"
author: "Jason Torres"
date: "July 26, 2016"
output: pdf_document
---

This script appends imputed models SNPs (elastic net alpha=0.5; GTEx Version 6p) to GWAS Z-score files for DIAGRAM (trans-ethnic) and GERA-T2D studies 


```{r setup, echo=FALSE}

library("dplyr")
library("data.table")
library("ggplot2")

"%&%" <- function(a,b) paste0(a,b)

server.dir <- "/Volumes/im-lab/nas40t2/jason/projects/MetaXcan/"

beta.dir <- server.dir %&% "db_files/gtex_v6p/alpha_0.5/allBetas/"

```



```{r get_all_model_snps, echo=FALS}

file.vec <- list.files(beta.dir)

get_all_mod_snps <- function(){
  snp.vec <- c()
  for (f in file.vec){
    print(f)
    df <- fread("cat " %&% beta.dir%&%f %&% " | zmore")
    snps <- df$rsid
    snp.vec <- append(snp.vec,snps)
  }
  return(unique(snp.vec))
}

snp.vec <- get_all_mod_snps()
saveRDS(snp.vec,file="/Volumes/im-lab/nas40t2/jason/projects/MetaXcan/" %&% "db_files/gtex_vp6/alpha_0.5/allBetas/all_tw_modSNPs.RDS")
```


# DIAGRAM (trans-ethnic) 

```{r diagram_setup, echo=FALSE}

# DIAGRAM (trans-ethnic) Build37
preimp.path <- server.dir %&% "meta_files/diagram/hm2ceu/diag3/diag3.txt.gz" 
postimp.path <- server.dir %&% "meta_files/" %&% "ImpGformat/merged_output_files/" %&% "impG_DIAGRAM/file/merged.imz.gz"

```



```{r evaluate_diagram_pre_imputation, echo=FALSE}

diag.df <- fread("cat " %&% preimp.path %&% " | zmore")
length(diag.df$SNP)
length(snp.vec)
print("There are " %&% length(intersect(diag.df$SNP,snp.vec)) %&% 
        " (" %&% round(length(intersect(diag.df$SNP,snp.vec))/length(snp.vec),3) %&% ")" %&% 
        " Model SNPs present in DIAGRAM (pre-imputation)")
```


```{r evaluate_diagram_post_imputation, echo=FALSE}

diag.post.df <- fread("cat " %&% postimp.path %&% " | zmore")
length(diag.post.df$SNP_id)
length(snp.vec)
print("There are " %&% length(intersect(diag.post.df$SNP_id,snp.vec)) %&% 
        " (" %&% round(length(intersect(diag.post.df$SNP_id,snp.vec))/length(snp.vec),3) %&% ")" %&% 
        " Model SNPs present in DIAGRAM (post-imputation)")
```

Now, will evaluate the additional set of Model SNPs introduced via imputation (ImpG-Summary)

```{r additional_mod_snps, echo=FALSE}

pre.mod.snps <- intersect(diag.df$SNP,snp.vec)
post.mod.snps <- intersect(diag.post.df$SNP_id,snp.vec)
addit.mod.snps <- setdiff(post.mod.snps,pre.mod.snps)
write.table(addit.mod.snps,file = server.dir %&% "meta_files/ImpGformat/merged_output_files/" %&% "impG_DIAGRAM/model-snps-gained_gtexV6p.txt",sep="\t",quote = FALSE, 
            col.names=FALSE,row.names=FALSE)

eval.df <- filter(diag.post.df,SNP_id %in% addit.mod.snps)


```

## Quality distribution 

* Prerequisite - Must run "00.2\_Append_r2pred.py script 

```{r quality, echo=FALSE}

file.name <- server.dir %&% "meta_files/ImpGformat/merged_output_files/impG_DIAGRAM/" %&% "model-snps-gained_gtexV6p_DIAGRAM.txt.gz"

df <- fread("cat " %&% file.name %&% " | zmore")
filt.df <- df[!duplicated(df),]

hist(filt.df$r2pred)
# proportion of additional imputed model SNPs meeting / not meeting quality threshold of 0.80 
sum(filt.df$r2pred >= 0.8) / dim(filt.df)[1]
sum(filt.df$r2pred <  0.8) / dim(filt.df)[1]


```


# GERA 

```{r gera_setup, echo=FALSE}

preimp.path <- server.dir %&% "meta_files/gera/gera/gera.txt.gz" 
postimp.path <- server.dir %&% "meta_files/ImpGformat/merged_output_files/impG_GERA/file/merged.imz.gz"

```


```{r evaluate_gera_pre_imputation, echo=FALSE}

gera.df <- fread("cat " %&% preimp.path %&% " | zmore")
length(gera.df$SNP)
length(snp.vec)
print("There are " %&% length(intersect(gera.df$SNP,snp.vec)) %&% 
        " (" %&% round(length(intersect(gera.df$SNP,snp.vec))/length(snp.vec),3) %&% ")" %&% 
        " Model SNPs present in GERA (pre-imputation)")
```


```{r evaluate_gera_post_imputation, echo=FALSE}

gera.post.df <- fread("cat " %&% postimp.path %&% " | zmore")
length(gera.post.df$SNP_id)
length(snp.vec)
print("There are " %&% length(intersect(gera.post.df$SNP_id,snp.vec)) %&% 
        " (" %&% round(length(intersect(gera.post.df$SNP_id,snp.vec))/length(snp.vec),3) %&% ")" %&% 
        " Model SNPs present in GERA (post-imputation)")
```

Now, will evaluate the additional set of Model SNPs introduced via imputation (ImpG-Summary)

```{r additional_mod_snps, echo=FALSE}

pre.mod.snps <- intersect(gera.df$SNP,snp.vec)
post.mod.snps <- intersect(gera.post.df$SNP_id,snp.vec)
addit.mod.snps <- setdiff(post.mod.snps,pre.mod.snps)
write.table(addit.mod.snps,file = server.dir %&% "meta_files/" %&% "ImpGformat/merged_output_files/" %&%  "impG_GERA/model-snps-gained_gtexV6p.txt",sep="\t",quote = FALSE, 
            col.names=FALSE,row.names=FALSE)

eval.df <- filter(gera.post.df,SNP_id %in% addit.mod.snps)


```


## Quality distribution 

* Prerequisite - Must run "00.2\_Append_r2pred.py script  

```{r quality, echo=FALSE}

file.name <- server.dir %&% "meta_files/ImpGformat/merged_output_files/" %&% "impG_GERA/model-snps-gained_gtexV6p_GERA.txt.gz"

df <- fread("cat " %&% file.name %&% " | zmore")
filt.df <- df[!duplicated(df),]

hist(filt.df$r2pred)
# proportion of addition imputed model SNPs meeting / not meeting quality threshold of 0.80 
sum(filt.df$r2pred >= 0.8) / dim(filt.df)[1]
sum(filt.df$r2pred <  0.8) / dim(filt.df)[1]


```

