---
title: "02.5_ComparePlots.Rmd"
author: "Jason Torres"
date: "August 4, 2016"
output:
  pdf_document:
    toc: yes
  html_document:
    fig_caption: yes
    highlight: textmate
    number_sections: yes
    theme: journal
    toc: yes
---


Here, I compare Significant associated gene sets from DIAGRAM (FDR 5% and Bonferonni) with significant gene sets from GERA-Replicaton (Pvalue < 0.05) 

*Analyses are limited to CrossTissue and Tissue-Wide models ONLY* 

# Setup

```{r setup, echo=FALSE}
library("dplyr")
library("ggplot2")
library("stargazer")
library("ggrepel")
"%&%" <- function(a,b) paste0(a,b) 
rds.dir <- "/Users/jtorres/Google Drive/Projects/MetaXcan/MetaXcanT2D/RDS.gtexV6p/" 
txt.dir <- "/Users/jtorres/Google Drive/Projects/MetaXcan/MetaXcanT2D/txt.gtexV6p/"
dir.create(txt.dir)
tab.dir <- "/Users/jtorres/Google Drive/Projects/MetaXcan/MetaXcanT2D/tables.gtexV6p/"
fig.dir <- "/Users/jtorres/Google Drive/Projects/MetaXcan/MetaXcanT2D/figures.gtexV6p/"
rep.df <- readRDS(rds.dir %&% "diag3/sig.genes.gera-rep.df.RDS")
```


```{r gera_threshold, echo=FALSE}

load(rds.dir%&%"gera/metaXcan.results.Rdata")
gera.thresh <- 0.05 /length(na.omit(as.numeric(as.matrix(select(p.EN.df,contains("TW_"))))))

```


# Z-score Plots of Replicated Gene-models pairs 

Directional comparison of Z-scores 

```{r zscore_plot}
zscore_plot <- function(alpha){
  rep.df$type <- as.factor(rep.df$type)
  rep.df$type <- factor(rep.df$type,levels(rep.df$type)[c(2,1,3)])
  labels <- c()
  for (i in 1:length(rep.df$gene)){
    gname <- rep.df$gene[i]
    repname <- rep.df$replicated[i]
    if (repname==TRUE){
      lname <- gname
    } else {
      lname <- ""
    }
    labels <- append(labels,lname)
  }
  rep.df <- cbind(rep.df,labels)
  col.vec <- c()
  for (i in 1:dim(rep.df)[1]){
    r = rep.df$replicated[i]
    p <- rep.df$gera.p.value[i]
    if (r==FALSE){
      col.vec<-append(col.vec,"grey")
    } else if(r==TRUE & p <= gera.thresh){
      col.vec<-append(col.vec,"dodgerblue4")
    } else{
      col.vec<-append(col.vec,"dodgerblue1")
    }
  }
  title <- "Directional Comparison between Studies" #\nalpha= " %&% alpha 
  plt <- ggplot(data=rep.df,aes(x=diag.z.score,y=gera.z.score)) +
    #geom_point(size=4,shape=21,aes(fill=replicated),show.legend=FALSE) +
    geom_point(size=2,shape=21,fill=col.vec,show.legend=FALSE) +
    geom_text_repel(data=filter(rep.df,replicated==TRUE,abs(diag.z.score)>=5,abs(gera.z.score)>=5,gera.p.value<=gera.thresh),aes(label=labels),
                    show.legend=FALSE,size=3,angle=30) + 
    geom_hline(yintercept=0) + geom_vline(xintercept=0) + 
    xlab("Z-score in DIAGRAM Study") + 
    ylab("Z-score in GERA-T2D Study") + 
    ggtitle(title) + 
    #scale_fill_brewer(name="Replicated",palette="Set3") + 
    scale_fill_manual(name="Replicated")+#,values=c("grey","dodgerblue1")) + 
    scale_x_continuous(breaks=seq(-10,10,2)) +
    scale_y_continuous(breaks=seq(-10,10,2))+
    scale_alpha_manual(values= c(0.5, 1)) + 
    theme_bw() + 
    theme(plot.title = element_text(lineheight=.8, face="bold"),
          axis.title.x = element_text(face="bold"),
          axis.title.y = element_text(face="bold"))
  outfile <- fig.dir %&% "GERA-Replication.Zscores.alpha" %&% alpha %&% ".pdf"
  ggsave(filename = outfile,plot=plt,device="pdf")
  plt
}
```


Directional comparison of Z-scores, restricted to predicted r2 >= 0.01 (1%) 


```{r zplot_1, echo=FALSE, fig.height=6,fig.width=9}
zscore_plot(0.5)
```

# SNP Number Plots 

```{r snp_used_plot, echo=FALSE}

#snp_used_plot <- function(alpha){
#  rep.df <- intersect_df(alpha)
#  rep.df$type.x <- as.factor(rep.df$type.x)
#  rep.df$type.x <- factor(rep.df$type.x,levels(rep.df$type.x)[c(2,1,3)])
#  title <- "Comparison of Number of Model SNP used between Studies\nalpha= " %&% alpha 
#  plt <- ggplot(data=rep.df,aes(x=num.used.snps.x,y=num.used.snps.y)) +
#    geom_point(shape=21,size=4,aes(fill=type.x)) + 
#    geom_text(aes(label=gene),nudge_y=0.5,size=3,angle=30) + 
#    #geom_hline(yintercept=0) + geom_vline(xintercept=0) + 
#    xlab("Number of Predictor SNPs in DIAGRAM3 Study") + 
#    ylab("Number of Predictor SNPs in GERA-T2D Study") + 
#    ggtitle(title) + 
#    scale_fill_brewer(name="Gene Class",palette="Set1") + 
#    theme_bw() + 
#    theme(plot.title = element_text(lineheight=.8, face="bold"),
#          axis.title.x = element_text(face="bold"),
#          axis.title.y = element_text(face="bold"))
#  outfile <- fig.dir %&% "GERA-Replication.NumSNPsUsed.alpha" %&% alpha %&% ".pdf"
#  ggsave(filename = outfile,plot=plt,device="pdf")
#  plt
#}
```


```{r splot_1, echo=FALSE, fig.height=6,fig.width=9}
#snp_used_plot(1)
```

# Create LaTeX output tables 

```{r table_function, echo=FALSE}

make.latex.table <- function(df,tabname,fname){
  outpath <- tab.dir %&% fname %&% ".tex"
  code <- stargazer(df,type="latex",title=tabname,summary=FALSE,
                    out=outpath,rownames=FALSE)
}
```

```{r make_rep_table_latex, echo=FALSE}

write_table <- function(alpha){
  #sub.df <- select(rep.df,chrom.x,gene,type.x,model.x,cv.r2.x,z.score.x,p.value.x,
  #                 z.score.y,p.value.y) %>% arrange(chrom.x)
  sub.df <- select(rep.df,gene,chrom,type,model,diag.z.score,diag.p.value,
                   gera.z.score,gera.p.value)
  sub.df$model <- gsub("TW_","",sub.df$model)
  names(sub.df) <- c("Gene","Chromosome","Class","Model","DIAGRAM Z-score",
                     "DIAGRAM p-value","GERA Z-score","GERA p-value")
  tabname <- "Summary of GERA-replicated genes" #", alpha = " %&% alpha 
  fname <- "GERA-Replication.table." %&% alpha 
  make.latex.table(sub.df,tabname,fname)
  ## save gene list 
  gene.list <- unique(filter(rep.df,replicated==TRUE)$gene)
  print("There are " %&% length(gene.list) %&% " unique genes")
  print("There are " %&% dim(filter(rep.df,replicated==TRUE))[1] %&% " associations that replicate in GERA")
  outname <- txt.dir %&% "GERA-Replication.genelist.alpha" %&% alpha %&% ".txt"
  write.table(gene.list,file=outname,quote=FALSE,sep="\t",row.names=FALSE,col.names=FALSE)
  return(sub.df)
}

tab.df <- write_table(0.5)

```



