---
title: "04.2_Get-MetaXcanSig-Windows.Rmd"
author: "Jason Torres"
date: "August 5, 2016"
output:
  pdf_document:
    toc: yes
  html_document:
    fig_caption: yes
    highlight: textmate
    number_sections: yes
    theme: journal
    toc: yes
---

In this script, I'll determine which non-overlapping windows have MetaXcan-significant associations. First, I'll determine those with genome-wide significant associations (i.e. bonferroni-significant across all tissue models). This will involve using the data frame res.df (processed version of T2Dcore.df). Second, we will need to iteratively determine windows that are locus-wide significant. 


```{r setup, echo=FALSE}

library("dplyr")

"%&%" <- function(a,b) paste0(a,b)

root.dir <- "/Users/jtorres/Google Drive/Projects/MetaXcan/MetaXcanT2D/"
rds.dir <- root.dir %&% "RDS/"
T2Dcore.df <- readRDS(rds.dir %&% "T2Dcore.df.RDS")

ref.df <- readRDS(rds.dir %&% "non.overlapping.t2dLoci.ref.df.RDS")
uniq.loci.df <- readRDS(rds.dir %&% "uniq.loci.df.RDS")

```


Determine non-overlapping genomic windows with Genome-wide (all tissue models) MetaXcan-significant associations in DIAGRAM (trans-ethnic) 

```{r gw_sig, echo=FALSE}

get_resdf <- function(){
  out.df <- T2Dcore.df
  out.df$diag.z.score <- as.numeric(out.df$diag.z.score)  
  out.df$diag.p.value <- as.numeric(out.df$diag.p.value)  
  out.df$gera.z.score <- as.numeric(out.df$gera.z.score)  
  out.df$gera.p.value <- as.numeric(out.df$gera.p.value) 
  out.df$meta.z.score <- as.numeric(out.df$meta.z.score)
  out.df$meta.p.value <- as.numeric(out.df$meta.p.value)
  return(out.df)
}

res.df <- get_resdf()

append_uni_sig <- function(){
  mtxn.sig.uni.genes <- c() 
  mtxn.sig.uni.models <- c() 
  ref.df$locus.genes <- as.character(ref.df$locus.genes)
  for (i in 1:dim(ref.df)[1]){
    gene.vec <- strsplit(ref.df$locus.genes[i],",")[[1]]
    temp.df <- filter(res.df,diag.bonfer==TRUE,gene %in% gene.vec)
    if (dim(temp.df)!=0){
      gstring <- paste(temp.df$gene,collapse=",")
      mstring <- paste(temp.df$model,collapse=",")
      mtxn.sig.uni.genes <- append(mtxn.sig.uni.genes,gstring) 
      mtxn.sig.uni.models <- append(mtxn.sig.uni.models,mstring) 
    } else{
      mtxn.sig.uni.genes <- append(mtxn.sig.uni.genes,NA) 
      mtxn.sig.uni.models <- append(mtxn.sig.uni.models,NA)       
    }
  }
  out.df <- cbind(ref.df, mtxn.sig.uni.genes, mtxn.sig.uni.models)
  return(out.df)
}

append_gw_sig <- function(){
  mtxn.sig.gw.genes <- c() 
  mtxn.sig.gw.models <- c() 
  ref.df$locus.genes <- as.character(ref.df$locus.genes)
  for (i in 1:dim(ref.df)[1]){
    gene.vec <- strsplit(ref.df$locus.genes[i],",")[[1]]
    #temp.df <- filter(res.df,diag.bonfer==TRUE,gene %in% gene.vec)
    temp.df <- filter(res.df,diag.mult.correct==TRUE,gene %in% gene.vec)
    if (dim(temp.df)!=0){
      gstring <- paste(temp.df$gene,collapse=",")
      mstring <- paste(temp.df$model,collapse=",")
      mtxn.sig.gw.genes <- append(mtxn.sig.gw.genes,gstring) 
      mtxn.sig.gw.models <- append(mtxn.sig.gw.models,mstring) 
    } else{
      mtxn.sig.gw.genes <- append(mtxn.sig.gw.genes,NA) 
      mtxn.sig.gw.models <- append(mtxn.sig.gw.models,NA)       
    }
  }
  #out.df <- cbind(ref.df, mtxn.sig.gw.genes, mtxn.sig.gw.models)
  out.df <- cbind(append_uni_sig(), mtxn.sig.gw.genes, mtxn.sig.gw.models)
  return(out.df)
}

```


Now, will determine the non-overlapping genomic windows with locus-wide MetaXcan associatons. 

```{r lw_sig, echo=FALSE}

load("RDS/diag3/metaXcan.results.Rdata")

append_lw_sig <- function(){
  ref.gw.df <- append_gw_sig()
  mtxn.sig.lw.genes <- c() 
  mtxn.sig.lw.models <- c() 
  #ref.df$locus.genes <- as.character(ref.gw.df$locus.genes)
  ref.gw.df$locus.genes <- as.character(ref.gw.df$locus.genes)

  for (i in 1:dim(ref.gw.df)[1]){
    gene.vec <- strsplit(ref.gw.df$locus.genes[i],",")[[1]]
    temp.df <- filter(p.L.df,Gene %in% gene.vec)
    p.thresh <- 0.05 / length(na.omit(as.numeric(as.matrix(temp.df[,2:dim(temp.df)[2]]))))
    gene.vec <- temp.df$Gene; mod.vec <- names(temp.df)[2:dim(temp.df)[2]]
    track.genes <- c()
    track.mods <- c()
    for (g in gene.vec){
      for (m in mod.vec){
        p.val <- as.numeric(filter(temp.df,Gene==g) %>% select(one_of(m)))
        if (!is.na(p.val) & p.val <= p.thresh){
          track.genes <- append(track.genes,g)
          track.mods <- append(track.mods, m)
        }
      }
    }
    if (length(track.genes)!=0){
      gstring <- paste(track.genes,collapse=",")
      mstring <- paste(track.mods,collapse=",")
      mtxn.sig.lw.genes <- append(mtxn.sig.lw.genes,gstring) 
      mtxn.sig.lw.models <- append(mtxn.sig.lw.models,mstring) 
    } else{
      mtxn.sig.lw.genes <- append(mtxn.sig.lw.genes,NA) 
      mtxn.sig.lw.models <- append(mtxn.sig.lw.models,NA)       
    }
  }
  out.df <- cbind(ref.gw.df, mtxn.sig.lw.genes, mtxn.sig.lw.models)
  return(out.df)
}


mtxn.sig.loci.df <- append_lw_sig()
saveRDS(object=mtxn.sig.loci.df,file=rds.dir%&%"mtxn.sig.loci.df.RDS")
mtxn.sig.loci.df <- readRDS(file=rds.dir%&%"mtxn.sig.loci.df.RDS")

```

There are `r dim(mtxn.sig.loci.df)[1]` non-overlapping T2D loci, of which:

`r length(na.omit(mtxn.sig.loci.df$mtxn.sig.uni.genes))` have genome-wide significant (across single model) MetaXcan associations 

`r length(na.omit(mtxn.sig.loci.df$mtxn.sig.gw.genes))` have genome-wide significant (across all models) MetaXcan associations 

`r length(na.omit(mtxn.sig.loci.df$mtxn.sig.lw.genes))` have locus-wide significant (across all models) MetaXcan associations 



