---
title: "05.2_GSEA-DIAGRAM.Rmd"
author: "Jason Torres"
date: "May 17, 2015"
output:
  pdf_document:
    toc: yes
  html_document:
    fig_caption: yes
    highlight: textmate
    number_sections: yes
    theme: journal
    toc: yes
---

**Gene Ontology GSEA analysis with goseq (Bioconductor)** 

# Setup 

```{r setup, echo=FALSE}
library("data.table")
library("dplyr")
library("ggplot2")
library("qvalue")
"%&%" <- function(a,b) paste0(a,b) 
rds.dir <- "intermediate_files/" 
txt.dir <- "intermediate_files/diag/" 
core.df <- readRDS(paste0(rds.dir,"diag/core.df.RDS"))
ggd.ref.df <- readRDS(paste0(rds.dir,"diag/ggd.ref.df.RDS"))
```

Genesets 

```{r genesets, echo=FALSE}

fdr.df <- readRDS(rds.dir%&%"FDR05.sig.df.RDS")
geneset.t2d <- filter(fdr.df, type=="T2D.Gene" | nhgri.gene == TRUE)$gene #readRDS(rds.dir%&%"replicated.p05.t2d.genes")
geneset.novel <- setdiff(fdr.df$gene, geneset.t2d) #readRDS(rds.dir%&%"replicated.p05.novel.genes")
geneset <- c(geneset.t2d,geneset.novel)

get_ensid <- function(genename){
  ensid <- as.character((filter(ggd.ref.df,symbol==genename) %>% dplyr::select(ensgene))[1,1]) 
  if (!is.na(ensid) & ensid=="character(0)"){
    ensid <- NA 
  }
  return(ensid) 
}

make_ensid_vec <- function(genevec){
  out.vec <- c()
  for (g in genevec){
    ensid <- get_ensid(g)
    out.vec <- append(out.vec,ensid)
  }
  return(out.vec)
}

```



**Dependencies**

NOTE: The following Code in the "Dependencies" and "Download the Data" section is from Jeff Leek's Statistics of Genome Sciences Coursera course lecuture on GSEA using goseq 

This document depends on the following packages:

```{r load_hidden, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library("data.table")
  library("dplyr")
  library("ggplot2")
  library(devtools)
  library(Biobase)
  library(goseq)
  library(DESeq2)
})
```

```{r load}
  library(devtools)
  library(Biobase)
  library(goseq)
  library(DESeq2)
```


To install these packages you can use the code (or if you are compiling the document, remove the `eval=FALSE` from the chunk.)

```{r install_packages, eval=FALSE}
#install.packages(c("devtools","MatrixEQTL"))
#source("http://www.bioconductor.org/biocLite.R")
#biocLite(c("Biobase","goseq","DESeq2"))
```

Download the data

Here we are going to follow along with the tutorial on [goseq](http://www.bioconductor.org/packages/release/bioc/vignettes/goseq/inst/doc/goseq.pdf). You can see what genomes are supported by this package


# Make input files for GSEA analyses   

Note: the vector files that will be created here will be further modifed in the next section within the GSEA functions 

First, need to create the input named vector, the vector will be integer indicating if a gene is in a significant gene set  

```{r gsea_inputs, echo=FALSE}

make_gsea_inputs <- function(geneset,prefix){
  analysis.genes <- unique(core.df$gene)
  ens.df <- filter(ggd.ref.df,symbol %in% analysis.genes,biotype=="protein_coding") %>% dplyr::select(symbol,ensgene)
  ens.vec <- c()
  for (g in analysis.genes){
    ens <- as.character(filter(ens.df,symbol==g)[1,2])
    ens.vec <- append(ens.vec,ens)
  }
  ens.vec <- unique(ens.vec)
  sig.ens.vec  <- na.omit(make_ensid_vec(geneset))
  int.vec <- c()
  for (e in ens.vec){
    if (e %in% sig.ens.vec){
      i <- 1
    } else{
      i <- 0 
    }
    int.vec <- append(int.vec,i)
  }
  sig.int.vec <- int.vec
  names(sig.int.vec) <- ens.vec
  sig.int.vec <- sig.int.vec[!duplicated(names(sig.int.vec))]
  saveRDS(sig.int.vec,paste0(rds.dir,prefix%&%".int.vec.RDS"))
  print("Saved RDS file")
}
```

# Gene set enrichment analyses (GSEA) Functions

Next, we will run a parametric GSEA analysis with go_seq  
We need to set up a weighting function for all the genes in that genome

Here we use a parametric test to look for differences in enrichment with respect to different categories. This is the 2 x 2 table approach. You can also use random sampling to define the null distribution by setting the parameters `method="Sampling"` and `repcnt=1000` for the number of null repititions. 

```{r gsea, echo=FALSE}

gsea <- function(named.int.vec,savename){
  analysis.genes <- unique(core.df$gene)
  ggd.sub.df <- ggd.ref.df; ggd.sub.df$chr <- as.numeric(ggd.sub.df$chr)
  total.genes <- filter(ggd.sub.df,biotype=="protein_coding",is.na(chr)==FALSE,symbol %in% analysis.genes) %>% dplyr::select(ensgene)
  total.genes <- unique(as.data.frame(total.genes,stringsAsFactors=FALSE)[,1])
  sig <- as.integer(total.genes %in% names(named.int.vec[named.int.vec == 1]))
  names(sig) <- total.genes
  pwf=nullp(sig,"hg19","ensGene")
  GO.wall=goseq(pwf,"hg19","ensGene")
  saveRDS(GO.wall,paste0(rds.dir,savename))
  print("Saved RDS file") 
}
```

You can also use random sampling to define the null distribution by setting the parameters `method="Sampling"` and `repcnt=1000` for the number of null repititions. (Jeff Leek) 

```{r gsea_nonparam, echo=FALSE}
gsea_nonparam <- function(named.int.vec,savename){
  analysis.genes <- unique(core.df$gene)
  ggd.sub.df <- ggd.ref.df; ggd.sub.df$chr <- as.numeric(ggd.sub.df$chr)
  total.genes <- filter(ggd.sub.df,biotype=="protein_coding",is.na(chr)==FALSE,symbol %in% analysis.genes) %>% dplyr::select(ensgene)
  total.genes <- unique(as.data.frame(total.genes,stringsAsFactors=FALSE)[,1])
  sig <- as.integer(total.genes %in% names(named.int.vec[named.int.vec == 1]))
  names(sig) <- total.genes
  pwf=nullp(sig,"hg19","ensGene")
  GO.Samp=goseq(pwf,"hg19","ensGene",method="Sampling",repcnt=10000) 
  saveRDS(GO.Samp,paste0(rds.dir,savename))
  print("Saved RDS file") 
}
```


```{r run_analyses, echo=FALSE}

append_qvals <- function(df){
  df <- filter(df,ontology=="BP")
  num <- dim(df)[1]
  bonf <- 0.05/num
  p.vec <- df$over_represented_pvalue
  qval <- qvalue(p.vec)$qvalues
  bonfer <- p.vec <= bonf
  out.df <- cbind(df,qval,bonfer)
  return(out.df)
}

make_gsea_inputs(geneset,"p05")
int.vec <- readRDS(rds.dir%&%"p05.int.vec.RDS")
gsea_nonparam(int.vec,"gsea.p05.df.RDS")
gsea.df <- readRDS(rds.dir%&%"gsea.p05.df.RDS")
gsea.df <- append_qvals(gsea.df)
saveRDS(gsea.df, rds.dir%&%"gsea.p05.df.RDS")

make_gsea_inputs(geneset.t2d,"p05.t2d")
int.vec <- readRDS(rds.dir%&%"p05.t2d.int.vec.RDS")
gsea_nonparam(int.vec,"gsea.p05.t2d.df.RDS")
gsea.t2d.df <- readRDS(rds.dir%&%"gsea.p05.t2d.df.RDS")
gsea.t2d.df <- append_qvals(gsea.t2d.df)
saveRDS(gsea.t2d.df, rds.dir%&%"gsea.p05.t2d.df.RDS")


make_gsea_inputs(geneset.novel,"p05.novel")
int.vec <- readRDS(rds.dir%&%"p05.novel.int.vec.RDS")
gsea_nonparam(int.vec,"gsea.p05.novel.df.RDS")
gsea.novel.df <- readRDS(rds.dir%&%"gsea.p05.novel.df.RDS")
gsea.novel.df <- append_qvals(gsea.novel.df)
saveRDS(gsea.novel.df, rds.dir%&%"gsea.p05.novel.df.RDS")


```


# Write tables 

```{r save_tables, echo=FALSE}

library("stargazer")
make.latex.table <- function(df,tabname,fname){
  outpath <- "intermediate_files/" %&% fname %&% ".tex"
  code <- stargazer(df,type="latex",title=tabname,summary=FALSE,
                    out=outpath,rownames=FALSE)
}

make.gsea.table <- function(df,extra=NULL){
  df <- filter(df,ontology=="BP")
  df <- dplyr::select(df,term,over_represented_pvalue)[1:20,]
  Rank <- 1:20; df <- cbind(Rank,df)
  names(df) <- c("Rank","GO:BP Pathway","P-value")
  tabname <- "Top GSEA Pathways for " %&% "FDR <= 0.05" %&% 
             "-significant genes from " %&% "DIAGRAM" %&% 
             " analysis" %&% extra
  fname <- "diag" %&% extra %&% ".gsea" 
  make.latex.table(df,tabname,fname)
}

make.gsea.table(gsea.df,extra="All")
make.gsea.table(gsea.t2d.df,extra="t2d")
make.gsea.table(gsea.novel.df,extra="novel")

```


