---
title: "GERA-StringentReplication"
author: "Jason Torres"
date: "August 15, 2016"
output: pdf_document
---

Here, I first consider all MetaXcan Genome-wide associations discoverd in DIAGRAM (trans-ethnic) and determine how they replicate in GERA at more stringent significance thresholds:
  * Bonferonni 
  * FDR 5% 
  * FDR 10% 
  

```{r setup, echo=FALSE}
"%&%" <- function(a,b) paste0(a,b)
library("dplyr")
library("data.table")
library("qvalue")
work.dir <- "/Users/jtorres/Google Drive/Projects/MetaXcan/MetaXcanT2D/"
rds.dir <- work.dir %&% "RDS/" 
```


Identify genome-wide significant associations in DIAGRAM 

```{r diagram, echo=FALSE}
get_gw_sig <- function(){
  load(rds.dir%&%"diag3/metaXcan.results.Rdata")
  gene.vec <- p.L.df$Gene
  mod.vec <- names(p.L.df)[2:length(names(p.L.df))]
  p.mat <- as.matrix(p.L.df[,2:dim(p.L.df)[2]])
  num.tests <- length(na.omit(as.numeric(p.mat)))
  gw.thresh <- 0.05 / num.tests
  sig.genes <- c()
  sig.mods <- c() 
  pvals <- c()
  for (g in 1:length(gene.vec)){
    for (m in 1:length(mod.vec)){
      gene <- gene.vec[g]
      mod <- mod.vec[m]
      pval <- p.mat[g,m]
      if (!is.na(pval) & pval <= gw.thresh){
        sig.genes <- append(sig.genes,gene)
        sig.mods <- append(sig.mods,mod)
        pvals <- append(pvals,pval)
      }
    }
  }
  out.df <- cbind(sig.genes,sig.mods,pvals)
  out.df <- as.data.frame(out.df)
  names(out.df) <- c("gene","model","p.value")
  out.df$gene <- as.character(out.df$gene)
  out.df$model <- as.character(out.df$model)
  out.df$p.value <- as.numeric(as.character(out.df$p.value))
  return(out.df)
}
gw.diag.df <- get_gw_sig()
```

Determine which associations replicate in GERA at genome-wide signficance, pvalue < 0.05, and fdr05,10,20% in a given tissue. 

```{r gera_gw, ehco=FALSE}

append_gera_p_and_gw <- function(df){
  # df must be in the format as output from get_gw_sig() 
  load(rds.dir%&%"gera/metaXcan.results.Rdata")
  gene.vec <- p.L.df$Gene
  mod.vec <- names(p.L.df)[2:length(names(p.L.df))]
  p.mat <- as.matrix(p.L.df[,2:dim(p.L.df)[2]])
  num.tests <- length(na.omit(as.numeric(p.mat)))
  gw.thresh <- 0.05 / num.tests
  gera.p <- c() 
  gera.gw.rep <- c()
  gera.fdr05.rep <- c()
  gera.fdr10.rep <- c()
  gera.fdr20.rep <- c()
  gera.p05.rep <- c() 
  for (i in 1:dim(df)[1]){
    gene <- df[i,1]
    mod <- df[i,2]
    p.df <- filter(p.L.df,Gene==gene) %>% select(one_of(mod))
    pval <- as.numeric(as.matrix(p.df))
    q.df <- filter(q.L.df,Gene==gene) %>% select(one_of(mod))
    qval <- as.numeric(as.matrix(q.df))
    if (!is.na(pval) & pval <= gw.thresh){
      rep <- TRUE 
    } else{
      rep <- FALSE
    }
    if (!is.na(pval) & pval <= 0.05){
      p05 <- TRUE 
    } else{
      p05 <- FALSE
    }
    if (!is.na(qval) & qval <= 0.05){
      fdr05 <- TRUE 
    } else{
      fdr05 <- FALSE
    } 
    if (!is.na(qval) & qval <= 0.10){
      fdr10 <- TRUE 
    } else{
      fdr10 <- FALSE
    }
    if (!is.na(qval) & qval <= 0.20){
      fdr20 <- TRUE 
    } else{
      fdr20 <- FALSE
    }
    gera.p <- append(gera.p,pval)
    gera.gw.rep <- append(gera.gw.rep,rep)
    gera.fdr05.rep <- append(gera.fdr05.rep,fdr05)
    gera.fdr10.rep <- append(gera.fdr10.rep,fdr10)
    gera.fdr20.rep <- append(gera.fdr20.rep,fdr20)
    gera.p05.rep <- append(gera.p05.rep,p05)
  }
  row.names(df) <- 1:dim(df)[1]
  out.df <- cbind(df,gera.p,gera.gw.rep,gera.fdr05.rep,
                  gera.fdr10.rep,gera.fdr20.rep,gera.p05.rep)#
  return(out.df)
}
gw.rep.df <- append_gera_p_and_gw(get_gw_sig()) 
saveRDS(gw.rep.df,file=rds.dir%&%"gw.rep.df.RDS")

```



# Alternate approach to q-values 


```{r alt_q, echo=FALSE}

reformat_df_add_q <- function(dataset.name){
  if (dataset.name == "DIAGRAM"){
    load(rds.dir%&%"diag3/metaXcan.results.Rdata")
  } else if (dataset.name == "GERA"){
    load(rds.dir%&%"gera/metaXcan.results.Rdata")
  } else{
    print("dataset.name must be DIAGRAM or GERA")
  }
  p.mat <- as.matrix(p.L.df[,2:dim(p.L.df)[2]])
  gene <- c() 
  model <- c() 
  p.value <- c() 
  gene.vec <- p.df$Gene
  mod.vec <- names(p.df)[2:dim(p.df)[2]]
  for (e in 1:length(mod.vec)){
    print(mod.vec[e])
    for (i in 1:length(gene.vec)){
      #p <- as.numeric(filter(p.df,Gene==g) %>% select(one_of(m)))
      p <- p.mat[i,e]
      if (!is.na(p)){
        g <- gene.vec[i]
        m <- mod.vec[e]
        gene <- append(gene,g)
        model <- append(model,m)
        p.value <- append(p.value,p)
      }
    }
  }
  out.df <- as.data.frame(cbind(gene,model,p.value))
  out.df$gene <- as.character(gene)
  out.df$model <- as.character(model)
  out.df$p.value <- as.numeric(as.character(out.df$p.value))
  print("Appending q-values...")
  q.value <- qvalue(out.df$p.value)$qvalues
  row.names(out.df) <- 1:dim(out.df)[1]
  out.df <- cbind(out.df,q.value)
  out.df$q.value <- as.numeric(out.df$q.value)  
  return(out.df)
}

diag.reform.df <- reformat_df_add_q("DIAGRAM")
gera.reform.df <- reformat_df_add_q("GERA")

reform.df <- inner_join(diag.reform.df,gera.reform.df,by=c("gene","model"))
saveRDS(reform.df,file=rds.dir%&%"reform.df.RDS")
```

```{r get_sig_alt, echo=FALSE}

get_gw_thresh <- function(dataset){
  if (dataset == "DIAGRAM"){
    load(rds.dir%&%"diag3/metaXcan.results.Rdata")
  } else if (dataset == "GERA"){
    load(rds.dir%&%"gera/metaXcan.results.Rdata")
  } else{
    print("dataset must be DIAGRAM or GERA")
  }
  gene.vec <- p.L.df$Gene
  mod.vec <- names(p.L.df)[2:length(names(p.L.df))]
  p.mat <- as.matrix(p.L.df[,2:dim(p.L.df)[2]])
  num.tests <- length(na.omit(as.numeric(p.mat)))
  gw.thresh <- 0.05 / num.tests
  return(gw.thresh)
}

make_alt_gw_rep_file <- function(){
  bonfer <- get_gw_thresh("DIAGRAM")
  df <- readRDS(file=rds.dir%&%"reform.df.RDS")
  df <- filter(df,p.value.x <= bonfer)
  bonfer2 <- get_gw_thresh("GERA")
  gera.gw.rep <- df$p.value.y <= bonfer2
  gera.fdr05.rep <- df$q.value.y <= 0.05
  gera.fdr10.rep <- df$q.value.y <= 0.10
  gera.fdr20.rep <- df$q.value.y <= 0.20
  gera.p05.rep <- df$p.value.y <= 0.05
  out.df <- cbind(df,gera.gw.rep,gera.fdr05.rep,gera.fdr10.rep,
                  gera.fdr20.rep,gera.p05.rep)
  return(out.df)
}

gw.rep_alt.df <- make_alt_gw_rep_file()
saveRDS(gw.rep_alt.df,file=rds.dir%&%"gw.rep_alt.df.RDS")
```


**NOTE**: The alternative approach, though more stringent, yields the same number of associations that replicate at FDR<=5% (11), and two less associations at FDR<=10% (12 vs 14; lose CAMK1D and JAZF1) and FDR<=20% (17 vs 19; lose HHEX, HMG20A, RCCD1), respectively.  

The 7 genes that replicate at FDR<=5%: "GPAM"    "HOXA5"   "PLEKHS1" "PPARG"   "SOX4"    "TCF7L2"  "WFS1"

The 4 genes that replicate at genome-wide signficance: "GPAM"    "HOXA5"   "PLEKHS1" "TCF7L2"



# Locus-wide replication

Here, I will determine if locus-wide significant associations in **DIAGRAM** (previously identified in 04 scripts) replicate in **GERA**


```{r setup_lw, echo=FALSE}

get_msig_df <- function(){
  df <- readRDS(rds.dir %&% "msig.loci.df.RDS")
  i <- sapply(df, is.factor)
  df[i] <- lapply(df[i], as.character)
  return(df) 
}

# This function will iterate through each of the 57 locus regions, extract the diagram p.value for each LW-association and then use the full set of locus genes, dplyr, and the p data frame in gera to determine if the association replicates at LW significance 

make_gera_lw_rep_df <- function(){
  df <- get_msig_df()
  load(rds.dir%&%"diag3/metaXcan.results.Rdata")
  diag.p.df <- p.L.df
  diag.z.df <- z.L.df 
  load(rds.dir%&%"gera/metaXcan.results.Rdata")
  gera.p.df <- p.L.df  
  gera.z.df <- z.L.df 
  rm(p.L.df,q.L.df,z.L.df)
  all.df <- readRDS(file=rds.dir%&%"reform.df.RDS")  
  gene <- c()
  model <- c()
  diag.z <- c()
  diag.p <- c() 
  gera.z <- c() 
  gera.p <- c()
  loc.thresh <- c()
  gera.locus.q <- c() 
  gera.strict.q <- c() 
  # Iterating through LOCI 
  for (i in 1:dim(df)[1]){
    print("Locus: " %&% i)
    temp.df <- filter(df,locus==i)
    loc.genes <- strsplit(temp.df$locus.genes,split=",")[[1]]
    lw.genes <- strsplit(temp.df$mtxn.sig.lw.genes,split=",")[[1]]
    lw.mods <- strsplit(temp.df$mtxn.sig.lw.models,split=",")[[1]]
    # check that there are lw genes at locus 
    if (!is.na(lw.genes)){
      for (e in 1:length(lw.genes)){
        g <- lw.genes[e]
        m <- lw.mods[e]
        dz <- as.numeric(filter(diag.z.df,Gene==g) %>% select(one_of(m)))
        dp <- as.numeric(filter(diag.p.df,Gene==g) %>% select(one_of(m)))
        gz <- as.numeric(filter(gera.z.df,Gene==g) %>% select(one_of(m)))
        gp <- as.numeric(filter(gera.p.df,Gene==g) %>% select(one_of(m)))
        p.vec <- na.omit(as.numeric(as.matrix(filter(gera.p.df,
                                              Gene %in% loc.genes)[,2:dim(gera.p.df)[2]])))
        num.tests <- length(p.vec)
        lt <- 0.05/num.tests
        # get locus-level q-value in GERA 
        q.vec <- qvalue(p.vec)$qvalues
        gq <- q.vec[match(gp,p.vec)]
        # look up strict (across all tests) GERA q value 
        gqs <- filter(all.df,gene==g,model==m)$q.value.y[1]
        # append values 
        gene <- append(gene,g)
        model <- append(model,m)
        diag.z <- append(diag.z,dz)
        diag.p <- append(diag.p,dp) 
        gera.z <- append(gera.z,gz) 
        gera.p <- append(gera.p,gp)
        loc.thresh <- append(loc.thresh,lt)
        gera.locus.q <- append(gera.locus.q,gq) 
        gera.strict.q <- append(gera.strict.q,gqs) 
      }
    }
  }
  out.df <- cbind(gene,model,diag.z,diag.p,
                  gera.z,gera.p,loc.thresh,gera.locus.q,gera.strict.q)
  out.df <- as.data.frame(out.df,stringsAsFactors=FALSE)
  out.df$diag.z <- as.numeric(out.df$diag.z)
  out.df$diag.p <- as.numeric(out.df$diag.p)
  out.df$gera.z <- as.numeric(out.df$gera.z)
  out.df$gera.p <- as.numeric(out.df$gera.p)
  out.df$loc.thresh <- as.numeric(out.df$loc.thresh)
  out.df$gera.locus.q <- as.numeric(out.df$gera.locus.q)
  out.df$gera.strict.q <- as.numeric(out.df$gera.strict.q)
  lw.rep <- out.df$gera.p <= out.df$loc.thresh
  lw.fdr05 <- out.df$gera.locus.q <= 0.05
  lw.fdr10 <- out.df$gera.locus.q <= 0.10
  lw.fdr20 <- out.df$gera.locus.q <= 0.20
  gw.fdr05 <- out.df$gera.strict.q <= 0.05
  gw.fdr10 <- out.df$gera.strict.q <= 0.10
  gw.fdr20 <- out.df$gera.strict.q <= 0.20
  gw.p05 <- out.df$gera.p <= 0.05 
  same.direct <- sign(out.df$diag.z) == sign(out.df$gera.z)
  out.df <- cbind(out.df, lw.rep, lw.fdr05, lw.fdr10, lw.fdr20,
                  gw.fdr05, gw.fdr10, gw.fdr20, gw.p05, same.direct)
  
}


rep.df <- make_gera_lw_rep_df()
saveRDS(rep.df,file=rds.dir%&%"lw.rep.df.RDS")


sum(na.omit(rep.df$lw.rep))
sum(na.omit(rep.df$lw.fdr05))
sum(na.omit(rep.df$lw.fdr10))
sum(na.omit(rep.df$lw.fdr20))
sum(na.omit(rep.df$gw.fdr05))
sum(na.omit(rep.df$gw.fdr10))
sum(na.omit(rep.df$gw.fdr20))
sum(na.omit(rep.df$gw.p05))

```


# Get replicated gene sets 

```{r rep_genes, echo=FALSE}

unique(filter(rep.df,lw.rep==TRUE)$gene)
unique(filter(rep.df,lw.rep==TRUE,same.direct==TRUE)$gene)

unique(filter(rep.df,lw.fdr05==TRUE)$gene)
unique(filter(rep.df,lw.fdr05==TRUE,same.direct==TRUE)$gene)

unique(filter(rep.df,lw.fdr10==TRUE)$gene)
unique(filter(rep.df,lw.fdr10==TRUE,same.direct==TRUE)$gene)

unique(filter(rep.df,lw.fdr20==TRUE)$gene)
unique(filter(rep.df,lw.fdr20==TRUE,same.direct==TRUE)$gene)

unique(filter(rep.df,gw.fdr05==TRUE)$gene)
unique(filter(rep.df,gw.fdr05==TRUE,same.direct==TRUE)$gene)

unique(filter(rep.df,gw.fdr10==TRUE)$gene)
unique(filter(rep.df,gw.fdr10==TRUE,same.direct==TRUE)$gene)

unique(filter(rep.df,gw.fdr20==TRUE)$gene)
unique(filter(rep.df,gw.fdr20==TRUE,same.direct==TRUE)$gene)

unique(filter(rep.df,gw.p05==TRUE)$gene)
unique(filter(rep.df,gw.p05==TRUE,same.direct==TRUE)$gene)

dim(filter(rep.df,gw.p05==TRUE))
dim(filter(rep.df,gw.p05==TRUE,same.direct==TRUE))
length(unique(filter(rep.df,gw.p05==TRUE)$gene))
length(unique(filter(rep.df,gw.p05==TRUE,same.direct==TRUE)$gene))


plot(rep.df$diag.z,rep.df$gera.z)
plot(filter(rep.df,same.direct==TRUE)$diag.z,filter(rep.df,same.direct==TRUE)$gera.z)


# Get Gene lists 
p05.genes <- unique(filter(rep.df,gw.p05==TRUE)$gene) # not-strict 
lw.fdr05.genes <- unique(filter(rep.df,lw.fdr05==TRUE)$gene) # stricter FDR <= 20% 
lw.fdr20.genes <- unique(filter(rep.df,lw.fdr20==TRUE)$gene) # more strict FDR <=5% 
lw.rep.genes <- unique(filter(rep.df,lw.rep==TRUE)$gene) # strictest; bonferonni 

# Collect reported genes 

reported.genes <- unique(readRDS(rds.dir %&% "loci.df.RDS")$reported.gene)
ebi.dir <- "/Volumes/im-lab/nas40t2/jason/projects/ReferenceData/NHGRI/"
ebi.file <- ebi.dir %&% "nhgri_ebi-trait-genes.txt.gz"
ebi.df <- fread("cat " %&% ebi.file %&% " | zmore")
names(ebi.df)<-c("trait","genes")
ebi.df <- filter(ebi.df,trait=="Type_2_diabetes")
ref.vec <- c()
for (r in ebi.df$genes){
  vec <- strsplit(r,",_")[[1]]
  ref.vec <- append(ref.vec,vec)
}
ref.vec <- unique(ref.vec)
reported.genes <- unique(append(reported.genes,ref.vec))

# Partition p05 genes 
p05.t2d.genes <- intersect(reported.genes,p05.genes)
p05.novel.genes <- setdiff(p05.genes,p05.t2d.genes)
saveRDS(p05.t2d.genes,rds.dir%&%"replicated.p05.t2d.genes")
saveRDS(p05.novel.genes,rds.dir%&%"replicated.p05.novel.genes")

# Partition lw.rep genes 
lw.rep.t2d.genes <- intersect(reported.genes,lw.rep.genes)
lw.rep.novel.genes <- setdiff(lw.rep.genes,lw.rep.t2d.genes) 

# Partition lw.fdr05 genes 
lw.fdr05.t2d.genes <- intersect(reported.genes,lw.fdr05.genes)
lw.fdr05.novel.genes <- setdiff(lw.fdr05.genes,lw.fdr05.t2d.genes) 

# Partition lw.fdr05 genes 
lw.fdr20.t2d.genes <- intersect(reported.genes,lw.fdr20.genes)
lw.fdr20.novel.genes <- setdiff(lw.fdr20.genes,lw.fdr20.t2d.genes) 

```




