---
title: "NHGRI-EBI_enrichement.Rmd"
author: "Jason Torres"
date: "October 13, 2016"
output: pdf_document
---

```{r setup,echo=FALSE}

library("data.table")
library("dplyr")
library("stargazer")

"%&%" <- function(a,b) paste0(a,b) 

rds.dir <- "intermediate_files/"
#rds.version1.dir <- "/Users/jtorres/Google Drive/Projects/MetaXcan/DIAGRAM-T2D/RDS/" 

nhgri.ebi.file <- "input_files/nhgri_ebi-trait-genes.txt.gz"

gwas.df <- fread("cat " %&% nhgri.ebi.file %&% " | zmore")
trait.vec <- unique(gwas.df$`DISEASE/TRAIT`); length(trait.vec)

tab.dir <- "intermediate_files/"

```


#Functions

```{r functions, echo=FALSE}

splitAndJoin <- function(vec){
  out.vec <- c() 
  vec <- strsplit(vec,",_")
  for (i in vec){
    for (e in i){
      out.vec <- append(out.vec,e)
    }
  }
  return(unique(out.vec))
}

nhgri_enrich_test <- function(sig.genes,n=1000,nhgri.genes){
  load(rds.dir%&%"diag/metaXcan.results.Rdata")
  gene.vec <- unique(p.EN.df$Gene)
  obs <- intersect(sig.genes,nhgri.genes)
  obs.count <- length(obs)
  print("Overalp is " %&% obs.count)
  perm.vec <- c()
  for (i in 1:n){
    samp <- sample(gene.vec,length(sig.genes),replace=FALSE)
    count <- length(intersect(samp,nhgri.genes))
    perm.vec <- append(perm.vec,count)
  }
  print("Max permuted value is " %&% max(perm.vec))
  emp.p <- (sum(perm.vec >= obs.count) + 1)  / (n + 1) # Added 1 to both numerator and denominator 
  #if (emp.p == 0){
  #  emp.p = 1/n
  #}
  return(emp.p) 
}

nhgri_enrich_test_outputList <- function(sig.genes,n=1000,nhgri.genes){
  load(rds.dir%&%"diag/metaXcan.results.Rdata")
  gene.vec <- unique(p.EN.df$Gene)
  obs <- intersect(sig.genes,nhgri.genes)
  obs.count <- length(obs)
  print("Overalp is " %&% obs.count)
  perm.vec <- c()
  for (i in 1:n){
    samp <- sample(gene.vec,length(sig.genes),replace=FALSE)
    count <- length(intersect(samp,nhgri.genes))
    perm.vec <- append(perm.vec,count)
  }
  print("Max permuted value is " %&% max(perm.vec))
  emp.p <- sum(perm.vec >= obs.count) / n
  if (emp.p == 0){
    emp.p = 1/n
  }
  return(list(obs.count,emp.p,perm.vec)) 
}

```


Determine which GWAS traits have reported genes enriched among FDR <= 0.05 significant MetaXcan results from DIAGRAM trans-ethnic meta-analysis 

```{r enrich_eval, echo=FALSE}



# less tests, n=100, to triage entire set of 1362 traits in gwas catalogue 
# and narrow down to a smaller list for more intensive testing 

triage_stage1 <- function(sig.genes){
  sigtraits.vec <- c() 
  count = 0 
  for (trait in trait.vec){
    count = count + 1 
    print ("Count: " %&% count)
    gene.vec <- unique(splitAndJoin(filter(gwas.df,`DISEASE/TRAIT`==trait)$`REPORTED_GENE(S)`))
    pval <- nhgri_enrich_test(sig.genes,n=100,nhgri.genes=gene.vec)
    if (pval <= 0.05){
      print(trait %&% ": " %&% pval)
      sigtraits.vec <- append(sigtraits.vec,trait)
    }
  }
  return(sigtraits.vec)
}
#t1.vec  <- triage_stage1()

triage_stage2 <- function(t1.vec,sig.genes){
  sigtraits.vec <- c()
  pval.vec <- c() 
  count = 0 
  for (trait in t1.vec){
    count = count + 1 
    print ("Count: " %&% count)
    gene.vec <- unique(splitAndJoin(filter(gwas.df,`DISEASE/TRAIT`==trait)$`REPORTED_GENE(S)`))
    pval <- nhgri_enrich_test(sig.genes,n=10000,nhgri.genes=gene.vec)
    if (pval <= (0.05/10)){
      print(trait %&% ": " %&% pval)
      #sigtraits.vec <- append(sigtraits.vec,trait)
    }
    sigtraits.vec <- append(sigtraits.vec,trait)
    pval.vec <- append(pval.vec,pval) 
  }
  return.df <- as.data.frame(cbind(sigtraits.vec,pval.vec),stringsAsFactors=FALSE)
  names(return.df) <- c("trait","p")
  return.df$p <- as.numeric(return.df$p)
  return(return.df)
}



```


All FDR05 significant genes

```{r fdr05_all, echo=FALSE}

fdr.df <- readRDS(file=rds.dir%&%"fdr05.processed.df.RDS")
fdr05.genes <- unique(fdr.df$gene)

#t1.vec  <- triage_stage1(fdr05.genes)
#sigtraits.df <- triage_stage2(t1.vec,fdr05.genes)
#sigtraits.df <- arrange(sigtraits.df,p)
#saveRDS(sigtraits.df,rds.dir %&% "diag/sigtraits.df.RDS")
#sigtraits.df <- readRDS(rds.dir %&% "diag/sigtraits.df.RDS")
```


All **NOVEL** FDR05 significant genes

```{r fdr05_novel, echo=FALSE}

fdr.df <- readRDS(file=rds.dir%&%"fdr05.processed.df.RDS")
fdr05.novel.genes <- unique(filter(fdr.df,type!="T2D.Gene")$gene)

t1.vec  <- triage_stage1(fdr05.novel.genes)
sigtraits.novel.df <- triage_stage2(t1.vec,fdr05.novel.genes)
sigtraits.novel.df <- arrange(sigtraits.novel.df,p)
saveRDS(sigtraits.novel.df,rds.dir %&% "diag/sigtraits.novel.df.RDS")
#sigtraits.novel.df <- readRDS(rds.dir %&% "diag/sigtraits.novel.df.RDS")
```


Known T2D genes among FDR05 significant genes

```{r fdr05_known, echo=FALSE}

fdr.df <- readRDS(file=rds.dir%&%"fdr05.processed.df.RDS")
fdr05.t2d.genes <- unique(filter(fdr.df,type=="T2D.Gene")$gene)

t1.vec  <- triage_stage1(fdr05.t2d.genes)
sigtraits.t2d.df <- triage_stage2(t1.vec,fdr05.t2d.genes)
sigtraits.t2d.df <- arrange(sigtraits.t2d.df,p)
saveRDS(sigtraits.t2d.df,rds.dir %&% "diag/sigtraits.t2d.df.RDS")
#sigtraits.t2d.df <- readRDS(rds.dir %&% "diag/sigtraits.t2d.df.RDS")
```





```{r save_latex_table, echo=FALSE}

get_overlapping_trait_genes <- function(mygene.vec,trait){
 gene.vec <- unique(splitAndJoin(filter(gwas.df,`DISEASE/TRAIT`==trait)$`REPORTED_GENE(S)`))
 overlap.vec <- unique(intersect(gene.vec,mygene.vec))#; print(overlap.vec)
 return(overlap.vec)
}

traits_linked_to_gene <- function(gene){
  out.vec <- c() 
  for (trait in trait.vec){
    gene.vec <- unique(splitAndJoin(filter(gwas.df,`DISEASE/TRAIT`==trait)$`REPORTED_GENE(S)`))
    if (gene %in% gene.vec){
      out.vec <- append(out.vec,trait)
    }
  }
  return(out.vec)
}

make.latex.table <- function(df,tabname,fname){
  outpath <- tab.dir %&% fname %&% ".tex"
  code <- stargazer(df,type="latex",title=tabname,summary=FALSE,
                    out=outpath,rownames=FALSE)
}


#reported.t2d.genes <- unique(c(diagram.reported.genes,nhgri.t2d.genes)) 

make_sigtrait_gene_df <- function(sigtraits.df,sig.genes){
  loci.df <- readRDS(rds.dir%&%"loci.df.RDS")
  diagram.reported.genes <- unique(loci.df$reported.gene)
  nhgri.t2d.genes <-  gene.vec <- unique(splitAndJoin(filter(gwas.df,`DISEASE/TRAIT`=="Type_2_diabetes")$`REPORTED_GENE(S)`))
  reported.t2d.genes <- unique(c(diagram.reported.genes,nhgri.t2d.genes)) 
  t.vec <- c()
  p.vec <- c()
  g1.vec <- c()
  g2.vec <- c()
  for (r in 1:dim(sigtraits.df)[1]){
    t <- sigtraits.df[r,1]
    p <- sigtraits.df[r,2]
    g <- get_overlapping_trait_genes(sig.genes, t)
    g1 <- intersect(reported.t2d.genes,g); g1 <- paste(g1,collapse=",") 
    g2 <- setdiff(g,reported.t2d.genes); g2 <- paste(g2,collapse=",")
    t.vec <- append(t.vec,t); p.vec <- append(p.vec,p)
    g1.vec <- append(g1.vec,g1); g2.vec <- append(g2.vec,g2)
  }
  df <- as.data.frame(cbind(t.vec,p.vec,g1.vec,g2.vec),stringsAsFactors=FALSE)
  names(df) <- c("t","p","g1","g2")
  df$p <- as.numeric(df$p)
  df$t <- gsub("_"," ",df$t)
  df <- filter(df,p <= 0.05) %>% arrange(p)
  names(df) <- c("Trait", "P-value", "Reported T2D Genes", "Novel Genes")
  return(df)
}

```


Write T2D genes (FDR <= 0.05) enrichment results

```{r write_1,echo=FALSE}

write.t2d.df <- make_sigtrait_gene_df(sigtraits.t2d.df,fdr05.t2d.genes)
make.latex.table(write.t2d.df,"nhgri-ebi enrichment; T2D genes","nhgri.enrichment.t2d")

```

Write Novel genes (FDR <= 0.05) enrichment results

```{r write_2,echo=FALSE}

write.novel.df <- make_sigtrait_gene_df(sigtraits.novel.df,fdr05.novel.genes)
make.latex.table(write.novel.df,"nhgri-ebi enrichment; Novel genes","nhgri.enrichment.novel")

```


# Determine what trait is linked to a particular gene 

```{r process, echo=FALSE}

test <- "TCF7L2"

t1 <- traits_linked_to_gene("UTS2"); print("UTS2"); t1
t2 <- traits_linked_to_gene("HLA-A");print("HLA-A"); t2
t3 <- traits_linked_to_gene("ZFAND6");print("ZFAND6"); t3
t4 <- traits_linked_to_gene("ZNRD1");print("ZNRD1");t4
t5 <- traits_linked_to_gene("FAH");print("FAH");t5

traits_linked_to_gene("TCF7L2")

```

# Shared etiology 

```{r shared, echo=FALSE}


get_overlapping_trait_genes(fdr05.genes, "Breast_cancer")
get_overlapping_trait_genes(fdr05.genes, "Breast_Cancer_in_BRCA1_mutation_carriers")
get_overlapping_trait_genes(fdr05.genes, "Glycated_hemoglobin_levels")
get_overlapping_trait_genes(fdr05.genes, "Proinsulin_levels")
get_overlapping_trait_genes(fdr05.genes, "LDL_cholesterol") 
get_overlapping_trait_genes(fdr05.genes, "Cholesterol,_total"); 
get_overlapping_trait_genes(fdr05.genes, "Triglycerides")
get_overlapping_trait_genes(fdr05.genes, "Body_mass_index")


```

